---
title: "p8105_hw6_rk3445"
author: "Rosie Kwon"
output: html_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
#Create a city_state variable, and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.
homicides = 
  read_csv("data/homicide-data.csv", na = c("NA", "Unknown", "")) |> 
  janitor::clean_names() |> 
  mutate(
    city_state = paste(city , state, sep = ", "),
    solve_status = case_match(
      disposition,
      "Closed without arrest" ~ 1,
      "Open/No arrest" ~ 1,
      "Closed by arrest" ~ 0
    ),
    victim_age = as.numeric(victim_age)
  ) |> 
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", 
                         "Kansas City, MO", "Tulsa, AL")),
    victim_race %in% c("White", "Black")
  ) 
```

The dataset from Washington Post homicide data contains information about murders in U.S. cities, including:

- `uid`: unique identifier for each case

- `victim_last`, `victim_first`: victim’s name

- `victim_age`, `victim_sex`, `victim_race`: demographics of victims

- `city`, `state`: location

- `disposition`: whether the case was solved or not

Dataset includes a total of `r nrow(homicides)` observations and `r ncol(homicides)` variables.
 
```{r}
#For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object;

homicides_md =
  homicides |> 
  filter(
    city_state == "Baltimore, MD"
  )

fit_md = 
  homicides_md |> 
  glm(
    solve_status ~ victim_age + victim_sex + victim_race, 
    family = binomial(link = "logit"), 
    data = _) 
```

```{r}
#apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
result =
  broom::tidy(fit_md, conf.int = TRUE, exponentiate = TRUE)

result |> 
  knitr::kable()
```

Since the results of a logistic regression model are given in log-odds, I took the exponential to convert the results to odd ratios (OR). 

```{r}
result |> 
  filter(term == "victim_sexMale") |> 
  select(term, estimate, conf.low, conf.high) |> 
  knitr::kable()
```

Male victims have a 2.35 times higher odds of having their homicide solved compared to female victims with holding other variables. The 95% confidence interval lies between 1.79 and 3.08.

```{r}
#Now run glm for each of the cities in your dataset
fit_cities = function(df){
  glm(
    solve_status ~ victim_age + victim_sex + victim_race, 
    family = binomial(link = "logit"), 
    data = df) 
}

#Extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

nested_results = 
  homicides |> 
  nest(data = -city_state) |> 
  mutate(
    fits = purrr::map(data, fit_cities),
    results = purrr::map(fits, ~broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))

  ) |> 
  select(city_state, results) |> 
  unnest(results) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, estimate, conf.low, conf.high) 


knitr::kable(nested_results)
```


```{r, fig.width= 12, fig.height= 25}
#Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

nested_results |> 
  mutate(city_state = fct_reorder(city_state, estimate)) |> 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  geom_hline(yintercept = 1, color = "red", linewidth = 0.4) +
  coord_flip() +
  scale_x_discrete(expand = expansion(add = 0.1)) +
  labs(
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female)",
    title = "Adjusted OR for Solving Homicides by Victim Sex"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12, margin = margin(r = 10))
  )
```

I added OR = 1 as a reference line to interpret which victim of sex is more likely to be solved. 

Most cities including New York, Baton Rouge, Cincinnati, Chicago, Long Beach, San Diego, and others show ORs above 1, indicating that male-victim homicides are more likely to be solved. Especially, New York City shows the largest ORs with wide confidence intervals, suggesting strong differences but also high uncertainty. 

When the 95% confidence interval includes 1, the difference is not statistically significant, meaning we cannot conclude that victim sex affects clearance rates in those cities. Importantly, although a few cities such as Albuquerque, Fresno and Stockton have ORs below 1, none of them show statistically significant evidence that cases involving female victims are more likely to be solved than male victims.

# Problem 2

```{r}
#Load library and central Park weather data
library(p8105.datasets)
library(modelr)
data("weather_df")
```

A simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data: Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities : $\hat{r}^2$, $\hat{\beta}_1 / \hat{\beta}_2$


```{r}
bootstrap_results =
  weather_df |> 
  bootstrap(n = 5000) |> 
  mutate(
    model = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    r_sq = map_dbl(model, \(m) broom::glance(m)$r.squared),
    beta_ratio = map_dbl(model, \(m) {
      tbl = broom::tidy(m)
      b1 = tbl$estimate[tbl$term == "tmin"]
      b2 = tbl$estimate[tbl$term == "prcp"]
      b1 / b2
    })
  ) |> 
  select(r_sq, beta_ratio)
```

Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for  $\hat{r}^2$, $\hat{\beta}_1 / \hat{\beta}_2$

1)  $\hat{r}^2$ estimate distribution
```{r}
bootstrap_results |> 
  ggplot(aes(x = r_sq))+
  geom_histogram() +
  labs(
    title = "Bootstrap Distribution of R²",
    x = "R²",
    y = "Count"
  ) +
  theme_minimal()
```

Using 5000 times of bootstrap, the distribution of the r_squared value is approximatlly normal, centered around its mean of `r mean(bootstrap_results$r_sq)`. Since r_squared is greater than 0.93, the power of the model is highly stable across resampled datasets.

```{r}
bootstrap_results |> 
  summarize(
    ci_lower = quantile(r_sq, 0.025), 
    ci_upper = quantile(r_sq, 0.975)
  ) |> 
  knitr::kable()
```

The 95% bootstrap confidence interval of $\hat{r}^2$ is between 0.934 and 0.946. 

2) $\hat{\beta}_1 / \hat{\beta}_2$ estimate distribution
```{r}
bootstrap_results |> 
  ggplot(aes(x = beta_ratio)) +
  geom_histogram() +
  labs(
    title = "Bootstrap Distribution of β1 / β2",
    x = "β1 / β2",
    y = "Count"
  ) +
  theme_minimal()
```

The ratio represents the change in `tmin` relative to `prcp` with the same change in `tmax`. The distribution of the ratio β1/β2 where β1 is estimate for `tmin` and β2 is estimate for `prcp`. It is left-skewed and wider and more variable because the precipitation coefficient is close to zero, making the ratio unstable and potentially very large in magnitude. 

```{r}
bootstrap_results |> 
  summarize(
    ci_lower = quantile(beta_ratio, 0.025), 
    ci_upper = quantile(beta_ratio, 0.975)
  ) |> 
  knitr::kable()
```

The 95% bootstrap confidence interval is between -280.71 and -125.20.

# Problem 3

Load and clean the data for regression analysis (i.e. use appropriate variable names, convert numeric to factor where appropriate, check for the presence of missing data, etc.).

```{r}
birthweight =
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace   = factor(frace, 
                     levels = c(1,2,3,4,8,9),
                     labels = c("White", "Black", "Asian", "Puerto Rican", 
                                "Other", "Unknown")),
    mrace   = factor(mrace,
                     levels = c(1,2,3,4,8),
                     labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform, levels = c(0,1), labels = c("No", "Yes"))
  )

birthweight |> skimr::skim()
```

The birthweight dataset has `r nrow(birthweight)` infants data with `r ncol(birthweight)` measurements. It is cleaned by standardizing variable names, converting appropriate numeric variables (infant sex, malformations, father’s race, mother’s race) into factors. I verified that the dataset doesn't contain any missing values. 

--

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. 

```{r}
linear_prop = lm(
    bwt ~ bhead + blength + gaweeks + ppbmi + smoken + wtgain + babysex, 
    data = birthweight)  
```

Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
birthweight |> 
  add_predictions(linear_prop) |> 
  add_residuals(linear_prop) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "red") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Birthweight",
    y = "Residual"
  ) +
  theme_minimal()

```


Compare your model to two others:

- A model using length at birth and gestational age as predictors (main effects only)

```{r}
linear_main = lm(bwt ~ blength + gaweeks, data = birthweight)  
```

- A model using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
linear_interaction = lm(bwt ~ bhead * blength * babysex, data = birthweight)
```


Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r}
cv_df = 
  crossv_mc(birthweight, n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_main = 
      map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    linear_interaction = 
      map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df)),
  ) |> 
  mutate(
    rmse_main = map2_dbl(linear_main, test, rmse),
    rmse_interaction = map2_dbl(linear_interaction, test, rmse)
  )
```

```{r}
cv_df |> 
  summarize(
    RMSE_Main   = mean(rmse_main),
    RMSE_Inter  = mean(rmse_interaction)
  ) |> 
  knitr::kable()
```

