---
title: "p8105_hw6_rk3445"
author: "Rosie Kwon"
output: html_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
#Create a city_state variable, and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.
homicides = 
  read_csv("data/homicide-data.csv", na = c("NA", "Unknown", "")) |> 
  janitor::clean_names() |> 
  mutate(
    city_state = paste(city , state, sep = ", "),
    solve_status = case_match(
      disposition,
      "Closed without arrest" ~ 1,
      "Open/No arrest" ~ 1,
      "Closed by arrest" ~ 0
    ),
    victim_age = as.numeric(victim_age)
  ) |> 
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", 
                         "Kansas City, MO", "Tulsa, AL")),
    victim_race %in% c("White", "Black")
  ) 
```

The dataset from Washington Post homicide data contains information about murders in U.S. cities, including:

- `uid`: unique identifier for each case

- `victim_last`, `victim_first`: victim’s name

- `victim_age`, `victim_sex`, `victim_race`: demographics of victims

- `city`, `state`: location

- `disposition`: whether the case was solved or not

Dataset includes a total of `r nrow(homicides)` observations and `r ncol(homicides)` variables.
 
```{r}
#For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object;

homicides_md =
  homicides |> 
  filter(
    city_state == "Baltimore, MD"
  )

fit_md = 
  homicides_md |> 
  glm(
    solve_status ~ victim_age + victim_sex + victim_race, 
    family = binomial(link = "logit"), 
    data = _) 
```

```{r}
#apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
result =
  broom::tidy(fit_md, conf.int = TRUE, exponentiate = TRUE)

result |> 
  knitr::kable()
```

Since the results of a logistic regression model are given in log-odds, I took the exponential to convert the results to odd ratios (OR). 

```{r}
result |> 
  filter(term == "victim_sexMale") |> 
  select(term, estimate, conf.low, conf.high) |> 
  knitr::kable()
```

Male victims have a 2.35 times higher odds of having their homicide solved compared to female victims with holding other variables. The 95% confidence interval lies between 1.79 and 3.08.

```{r}
#Now run glm for each of the cities in your dataset
fit_cities = function(df){
  glm(
    solve_status ~ victim_age + victim_sex + victim_race, 
    family = binomial(link = "logit"), 
    data = df) 
}

#Extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

nested_results = 
  homicides |> 
  nest(data = -city_state) |> 
  mutate(
    fits = purrr::map(data, fit_cities),
    results = purrr::map(fits, ~broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))

  ) |> 
  select(city_state, results) |> 
  unnest(results) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, estimate, conf.low, conf.high) 


knitr::kable(nested_results)
```


```{r, fig.width= 12, fig.height= 25}
#Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

nested_results |> 
  mutate(city_state = fct_reorder(city_state, estimate)) |> 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  geom_hline(yintercept = 1, color = "red", linewidth = 0.4) +
  coord_flip() +
  scale_x_discrete(expand = expansion(add = 0.1)) +
  labs(
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female)",
    title = "Adjusted OR for Solving Homicides by Victim Sex"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12, margin = margin(r = 10))
  )
```

I added OR = 1 as a reference line to interpret which victim of sex is more likely to be solved. 

Most cities including New York, Baton Rouge, Cincinnati, Chicago, Long Beach, San Diego, and others show ORs above 1, indicating that male-victim homicides are more likely to be solved. Especially, New York City shows the largest ORs with wide confidence intervals, suggesting strong differences but also high uncertainty. 

When the 95% confidence interval includes 1, the difference is not statistically significant, meaning we cannot conclude that victim sex affects clearance rates in those cities. Importantly, although a few cities such as Albuquerque, Fresno and Stockton have ORs below 1, none of them show statistically significant evidence that cases involving female victims are more likely to be solved.
